---
title: "Homework 3 -- Hypothesis testing basics"
output:
  html_document:
    df_print: paged
date: 'Due date: Tuesday September 24, 2019'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hypothesis testing by exact computations

Write down the null, the alternative, the test, and compute the p-value for each of the following scenarios. 

1. Amit claims that he can taste the difference between diet and regular soda. We gave Amit 10 cups to taste test, 5 of each type, but we did NOT tell Amit that there were 5 of each type. He got 8 cups right. Does the evidence support his claim? 

Null - Either Amit guess randomly or not.
Alternative -Amit can tell differnece between the reuglar coke and diet coke.
Test - The each of the tests are independent from each other. Each time you have 50% of pick right. We can test the hypothesis by test by using binomial distribution testing that he pick S>=8.


```{r}
 p.true.value1 <- (choose(10,8)*(.5)**8*(.5)**2) +(choose(10,9)*(.5)**9*(.5)**1) +(choose(10,10)*(.5)**10)
p.true.value1
```



2. Same as 1., but  we DID tell Amit that there were 5 of each type. (He still got 8 cups right). 

Null - Either Amit guess randomly or not.
Alternative - Amit can tell difference between the reuglar coke and diet coke .
Test -In the second one, you are constrained by the fact that you can only pick diet 5 times and regular 5 times during each sample. So the trials are not indenpendent.
we test find probability by 5 choose 4 right and 5 choose 5 right (S>=8) but pairs.
```{r}
p.true.value2 <- choose(5,4)^2/choose(10,5) + choose(5,5)^2/choose(10,5)
p.true.value2

```




3. Same as 1., but we did not know about Amit's claim, and simply wanted to test whether he was guessing randomly or not. (He still got 8 cups right). 

Null - Either Amit guess randomly or not.
Alternative - Amit's not guess randomly
Test -Caculate p- value for if Amit get right for  EXTREME(0,1,2,8,9,10)
```{r}
n = 10
p = .5
mean <- n*p
distribution.from.mean <- 8 - mean
p.true.value3 <- 1 -(pbinom(8-1, size =10, prob =0.5 )- pbinom(2,size =10, prob =0.5))
p.true.value3
```
## Simulating data under the null

4. For each of the above three scenarios, simulate 10 thousand datasets under the null hypothesis, and use your data to numerically approximate the $p$-value.

Null - Simulate 10 thousand datasets for each scenario to find p-value and see if Amit guess randomly.
Alternative - Amit's not guess random;y.
Test - Test 10,000 times for each senarios and recalculate the p-Value.
```{r}
num.exp=10**4
oneTrial <- rbinom(10* num.exp,1,0.5)
trialmatrix <- matrix(oneTrial,nrow=num.exp)

true.sequence <- c(1,1,1,1,1,0,0,0,0,0)
count.by.row <- function(row)
{
  return(sum(true.sequence==row))
}
S <- apply(trialmatrix,1, count.by.row)
p.hat1 <- sum(S>=8)/length(S)
p.value1 <- (abs(p.hat1 - p.true.value1))
p.hat1
```

```{r}
one.guess <- c("D","D","D","D","D","R","R","R","R","R")
num.exp <- 10**4
data <- matrix(replicate(num.exp,sample(one.guess)), ncol = 10,
               nrow = num.exp, byrow = T )
true.seq <- c("D","D","D","D","D","R","R","R","R","R")
count.correct <- function (row)
{
return(sum(row == true.seq))  
}
S <- apply(data, 1 , count.correct)
p.hat2<- sum(S>=8)/length(S)
p.hat2
```
```{r}
num.exp=10**4
oneTrial <- rbinom(10* num.exp,1,0.5)
trialmatrix <- matrix(oneTrial,nrow=num.exp)

true.sequence <- c(1,1,1,1,1,0,0,0,0,0)
count.by.row <- function(row)
{
  return(sum(true.sequence==row))
}
S <- apply(trialmatrix,1, count.by.row)
p.hat3 <- sum(S<=2,S>=8)/length(S)
p.hat3
```
We concludes that P values are small so it's more likely null hypothesis are true, but it could be wrong.
