---
title: 'Homework 6: approximate confidence intervals with the normal and t distribution'
output:
  html_document:
    df_print: paged
date: 'Due date: October 15, 2019'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We encourage you to use R as a calculator to solve these exercises. (As opposed to punching those numbers into your physical calculator). 

## Confidence intervals for proportions

Do textbook exercises 5.8, 5.12, 6.2, 6.8, 6.10,

5.8

Sample mean x = 0.52
crtical z for the CI  2.575829
standard error (SD) = 0.024

the 99% confidence interval for the US adult twitter users who get some news on twitter is 45.82% and 58.18%

5.12
The General Social Survey asked the question:for how many days during the past
30 days was your mental health, which includes stress, depression, and problems with emotions, not good?"
Based on responses from 1,151 US residents, the survey reported a 95% confidence interval of 3.40 to 4.24
days in 2010.
(a) Interpret this interval in context of the data.

95% of the multiple experiments has the interval at (3.40, 4.240)

(b) What does 95% confiden mean? Explain in the context of the application.

95% confident means that if the experiment procedue multiple times and the the there is a interval in that experiment. Let's say you wake up at 9:00 am or 9:10 am without any alarms. After many experiment this person conclude that 95% confidence interval of the wake up time is 9:00 am and 9:10am. 

(c)Suppose the researchers think a 99% confidence level would be more appropriate for this interval. Will
this new interval be smaller or wider than the 95% confidence interval
99% confidence interval in this will be larger than the 95% confidence interval.

(d)If a new survey were to be done with 500 Americans, do you think the standard error of the estimate
be larger, smaller, or about the same.
The standard error will be larger


6.2
a). True 
b). False since the the statment is for sample means but not for the sample proportions
c).
Let P(p hat>0.85) = 0.85 - 0.77 / (sqrt(.77(1-.77)/60))
      = 1 - P(z<= 1.47)
      = 1 - 0.-708
      = 0.0708
Which is greater than 0.05
there for violate the usual condition
Therefore statment is false.
d).
Let P(p hat>0.85) = 0.85 - 0.77 / (sqrt(.77(1-.77)/120))
                  = 1-P(z<= 2.08)
                  = 1-0.9812
                  = 0.0188
Which is less than 0.05
This satifies the usual condition
Therefore statment is ture
                  
6.8

a) Point estimate of parameter is 0.25
b) n p = 1000 * 0.25 = 250 > 10
c) 
a = 0.05 , z(0.025) = 1.96
z +- sqrt(p*(1-p)/n)
0.25 +- 1.96*sqrt(0.25*(1-0.25)/10000)
(0.22316,0.2768384)

d) if we deviceded to use a higher confidence level then confidence. interal gets larger.

e) if we use large sample size then interval decrease, which mean a more accuate esatimate of the population value.

6.10

a) 61% is a sample statistic since it is based on a sample of 1578 US resident.
b)
p hat = .61
sample size n = 1578
cofidence level 1 - a = 0.95
level of sigificance a = 0.05
p+- z sqrt(p(1-p)/n)
0.61+- 1.96sqrt(((.61)(1-0.61))/1578)
0.61+-0.02406577172
(0.58593422827,0.63406577172)
95% confident interval that US resident who think marijuana should be made legal is between 0.04524 and 0.5079

c)
np = 1578*.61 = 962.58
n(1-p) = 1578(1-.61) = 615.42
since both np and n(1-p) are least 10 and sample observations are indenpendent, the statistic sample proportion will be normally distributed.

d) since lower limit of confidence interval is less than 0.5, less than 50% population is also a plausible value of true proportion. So majority of American is not justified.
np


## Confidence intervals

Do textbook exercises 6.22, 6.30, 7.16, 7.18, 7.26
6.22

P(Z<=1.96) = 0.05
Confidence interval
=(0.08-0.088)+- 1.96(sqrt(0.08(1-.08)/11,545+(0.088(1-.088)/11,545))
=-0.008 +- 0.0095
= (-0.008-0.0095. 0.23 + 0.0386)
=(-0.0175,0.0015)

6.30
No, because it's same the same sample population and not indenpendent obersavation.

7.16

a) True
b) True
c) False, only if it was depdent sample
d) false, each observations are different.

7.18
a) paired
b) paired
c) Not paried



7.26
```{r}
summary <- c('Mean', 'SD', 'n')
auto <- c(44.51, 56.81, 23)
manual <- c(13.32, 16.13, 23)

df <- data.frame(summary,auto,manual); df
mean.diff <- df$auto[1] - df$manual[1]
se <- sqrt((((df$auto[2])^2/26) + ((df$m[2])^2/26)))
t.value <- mean.diff/se; t.value
p.value <- pt(t.value, df=25); p.value
```

## Various other problems

Do textbook exercises: 7.6, 7.14, 7.22, 7.28, 7.30

7.6

symmetic Confidenc interval
sample mean (x) = (65+77)/2 = 71
x = 71

Margin of error E = 1/2 x length of CI = (65-77)/2 = 6

The critical value t0.05 =1.711

E = t * s / sqrt(n)
s = E/z * sqrt(n)
s = 17.5336

7.14
standard devidation(s) = 250
margin of error  E =25
levvel of significance (a) = 0.10

n = ((Z x s) /E )^2
n = ((1.645 x250)/25)^2
  = (16.45)^2
  = 270.6
  = 271
b) if the confidence interval incresases, the sample size also increase.
  

7.22,
sample size N = 200
Mean  = -0.545
Standard Deviation = 8.887
x+- z * s/sqrt(n)
-0.545 +- 1.96*8.887/sqrt(200)
= (-1.777,0.687)

We can conclude that 95% cofident thatt average differences between the reading and writing scores of all student are in the interval of (-1.777,0.687)

c) The 95 % confidence interval contain zero value, so null hypothesis is not rejected. The confidence interval provides that there is not real difference in the avearage scores among students.




7.28,
```{r}
summary <- c('Mean', 'SD', 'n')
auto <- c(16.12, 3.58, 26)
manual <- c(19.85, 4.51, 26)

df <- data.frame(summary,auto,manual); df
mean.diff <- df$auto[1] - df$manual[1]
se <- sqrt((((df$auto[2])^2/26) + ((df$m[2])^2/26)))
t.value <- mean.diff/se; t.value
p.value <- pt(t.value, df=25); p.value
```
7.30
```{r}
summary <- c('Mean', 'SD', 'n')
auto <- c(22.92, 5.29, 26)
manual <- c(27.88, 5.01, 26)

df <- data.frame(summary,auto,manual); df
mean.diff <- df$auto[1] - df$manual[1]
se <- sqrt((((df$auto[2])^2/26) + ((df$m[2])^2/26)))
t.value <- mean.diff/se; t.value
p.value <- pt(t.value, df=25); p.value
```

 
