---
title: 'Homework 8: Linear regression basics'
output:
  html_document:
    df_print: paged
date: 'due date: October 29th, 2019'
---


## Regression basics

Textbook exercises: 8.6, 8.8

## Interpreting R's output of a regression model

In lecture 12, we investigated the relationship between math and read scores in the High School and Beyond survey. In this homework, you will take a closer look at the R's model output. 

1. Load the dataset \texttt{hsb2} in R from the library \texttt{openintro}. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openintro)

```

2. Produce a scatterplot of math vs read score. If we want to predict math scores based on read scores, which variabled should be on the $y$-axis?
```{r}
data(hsb2)
plot(hsb2$math ~ hsb2$read)
```

If we want to predict math scores based on read scores, math score should be on the $y$-axis, 

3. The following command fits a linear regression model $(math) = a +  b \times (read)$ to the data, and store the output in a variable called \texttt{model}

Run this command, print the model, and write down the equation that R fitted to the data.

(math)= 21.0382 + 0.6051 * (read)
```{r}
model <- lm(math ~ read, data =hsb2)
model
```

4. Interpret the intercept ($a$) and the slope ($b$) in the context of this problem (ie: what do these numbers say about read scores and math scores?)

$a$ - the average math score is 21.0382 when reading scores equal to 0.6051, .
$b$ - represent the change of the math scores as the reading score increase.

5. Adapt the code from lecture to plot the above equation on top of the scatterplot you had before. Based on this plot, do you think that the linear model is a good fit?
```{r}
data(hsb2)
plot(hsb2$math ~ hsb2$read)
hsb2.model = lm(math ~ read, data = hsb2)
abline(hsb2.model$coef, col="red")
```

constant variance indenpdent since evenlyy distribute in line and there is a trend in the line. Therefore I think linear model is a good fit.