---
title: "Homework 4"
output:
  html_document:
    df_print: paged
date: 'Due date: October 1, 2019'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (Tuesday HackR worksheet) The drawback of permutation tests

Permutation tests give a numerical procedure to approximate a $p$-value. Its main drawback is that it is computationally expensive: often one has to average over a large number of simulations to get a reliable estimate of the $p$-value. This exercise demonstrates this point. The bottom line is: if you knew an exact formula for the $p$-value, you should *always* use it.

Consider setup 2. of the soda taste test experiment from the last homework again. That is, the tester tasted 10 cups, knew that there were 5 of each type. Let $S$ be the number of cups that a taste tester got right under random guessing. 

1. Write a function so that it takes an argument $m$ and
  repeat the previous task $m$ times and return their outputs as a vector. That is, your function returns the output of $m$ independent simulations of $S$. Call your function \texttt{sodashuffle}.

#let 1 = "R" and 0 = "D"
```{r}
sodashuffle <- function(m){  
  
  one.guess <- c("D","D","D","D","D","R","R","R","R","R")
  num.exp <- m
  data <- matrix(replicate(num.exp,sample(one.guess)), ncol = 10,
               nrow = num.exp, byrow = T )
  true.seq <- c("D","D","D","D","D","R","R","R","R","R")
  count.correct <- function (row)
  {
  return(sum(row == true.seq))  
  }
  S <- apply(data, 1 , count.correct)
  S
}
sodashuffle(6)
```
  
2. Use \texttt{sodashuffle} to draw $n = 1000$ draws from the distribution of $S$. From this data, compute a numerical approximation $\hat{p}$ for $p := \mathbb{P}(S = 10)$. How far is this approximation from the true value of $p$, which is $1/252$? 
```{r}
p.hat.soda <- (sum(sodashuffle(1000) == 10))/1000
p.hat.soda
p.value.soda <- abs(((sum(sodashuffle(1000) == 10))/1000)-1/252)
p.value.soda
  
  
```
P value are very small.
3. Write a function to take $n$ (number of draws) as an argument, and return the error $|\hat{p} - p|$ when one performs step 2 with $n$ draws. 
```{r}
error <- function(n){abs(((sum(sodashuffle(n) == 10))/n)-1/252)}
error
```

4. Write a function to take $r$ (the number of runs) AND $n$ (number of draws) as arguments, and a vector obtained from repeating the function in step 3 $r$ times.  
```{r}
repeat.error <- function(n,r){replicate(r,abs(((sum(sodashuffle(n) == 10))/n)-1/252))}
repeat.error
```
  4. Plot your error distribution of $|\hat{p} - p|$ as a series of boxplots, one for each $n$, for $n = 2^5,2^6,2^7,2^8,2^9,2^{10},2^{11},2^{12},2^{13}$.
box
```{r}
error.matrix <- cbind(replicate(30,error(2**5),replicate(30,error(2**6),replicate(30,error(2**6),replicate(30,error(2**7),replicate(30,error(2**8),replicate(30,error(2**9),replicate(30,error(2**10),replicate(30,error(2**11),replicate(30,error(2**12),replicate(30,error(2**13))))))))))))
boxplot(error.matrix)

```
5. What can you conclude from this plot?
error distribution of $|\hat{p} - p|$ very row, so p hat values are close to p true value, which mean an approximation was pretty accurate.

## (Tuesday HackR worksheet) Titanic dataset. 
You can find this under homework4/titanic.csv. Read the file description from homework4/titanic-descrip.txt
Compare the survival rate between classes. Include an appropriate plot.
```{r,message=FALSE,fig.align='center',fig.show='hold',fig.width=3,fig.height=3}
library(ggplot2)
library(vcd)
library(openintro)
setwd("/Users/williamkwon/Documents/R project folder")
data <- read.csv("titanic.csv", header = T)
table(data$survived, data$class)
ggplot(data, aes(x=class,fill=survived)) + geom_bar()
```
crew(76%) and third class(death raito 75%) have less survival rates than first and second classes.

1. Is there a significant difference in survival rate between classes? 


```{r}
#used tabale and tab find the rates
counts <- c(325,285,706,885)
rates <- c(.24,.62,.41,.25)
survived.counts <- round(counts*rates)
not.survived.count <- counts-survived.counts
#table should have two variables: 
#class group (numbered 1 - 4) and use (=1 if they survived, 0 if not).
#matrices of titanic survival:
not.survived.list <- list()
for(i in c(1:4)){  
  not.survived.list[[i]] <- matrix(rep(c(i,0),not.survived.count[i]),ncol=2,byrow=T)
}
#matrices of non titatnic survival:
survived.list <- list()
for(i in c(1:4)){  
  survived.list[[i]] <- matrix(rep(c(i,1),survived.counts[i]),ncol=2,byrow=T)
}
#stack the individual matrices:
data <- not.survived.list[[1]]
for(i in c(2:4)){  
  data <- rbind(data, not.survived.list[[i]])
}
for(i in c(1:4)){  
  data <- rbind(data, survived.list[[i]])
}
```
```{r}
data <- as.data.frame(data)
names(data) <- c("class.group","survived")
data
data.deviation <- function(data){  
   
  survived.by.class <- by(data$survived,data$class,sum)  
  #turn this into a rate  
  rate.by.class <- survived.by.class/counts  
  #compute deviation   
  average.rate <- sum(rate.by.class)/sum(counts)  
  deviation <- sum(abs(rate.by.class-average.rate))  
  return(deviation)
}


```
```{r}
shuffle <- function(){  
  survived.shuffle <- sample(data$survived)  
  #compute total use by age from the shuffled use  
  survived.by.class <- by(survived.shuffle,data$class.group,sum)  
  #turn this into a rate  
  rate.by.survived <- survived.by.class/counts  
  #compute deviation   
  average.rate <- sum(survived.by.class)/sum(counts)  
  deviation <- sum(abs(rate.by.survived-average.rate))  
  return(deviation)
}
```
There is a significant difference in survival rate between classes since p-value equal to zero. For example crew(76%) and third class(death raito 75%) have less survival rates that first and second classes.

### Is there a significant difference in survival rate between classes? 

2. Set up the problem as a hypothesis testing problem. State the Null, the Alternative, and the event E
Null - There is no  difference in survive rate by the class.
Alternative - There is a difference in survive rate by the class.
Event - A good test is measured by the value:
\[ D = \sum_{i} |d_i - \overline{d}|  \]
where $d_i$ are the percentages of survival rate in each classes. This measures the deviation from the population average. $p$-value: Using the test statistic above, the $p$-value is:\[ p  = \mathbb{P}(D \geq d) \] - We can use the permutation test to compute an approximation to $p$. Then use $p$ to test Null hypothesis.


3. Compute an estimation of the p-value using the *permutation test* (*NO* other tests accepted).
```{r}
m = 10**4
D = replicate(m,shuffle())
#calculate the actual data deviation
d = data.deviation(data)
p.value <- sum(D >= d)/m
p.value
```
There is a significant difference in survival rate between classes since p-value equal to zero. For example crew(76%) and third class(death raito 75%) have less survival rates that first and second classes.





4. Draw conclusions. 

The calculated p value is 0, so we should reject $H_0$. Therefore there is a significant difference in survival rate between classes. 


## Teacher dataset. 
Consider the dataset ```teacher``` in the library ```OIdata```. This data set contains teacher salaries from 2009-2010 for 71 teachers employed by the St. Louis Public School in Michigan, as well as several other variables. Load the data in R with 
```{r}
library(OIdata)
data(teacher)
```
for documentation, type 
```{r}
?teacher
```
1. Compare the distribution of base salary in *full time* teachers with BA vs MA. Include an appropriate plot. 
```{r,message=FALSE,fig.align='center',fig.show='hold',fig.width=3,fig.height=3}
fulltime <- subset(teacher,fte==1) 
table(fulltime$degree, fulltime$base)
ggplot(fulltime, aes(x= degree,fill=base)) + geom_bar()
```

### Do *full time* teachers with an MA start with a significantly higher *minimum* base salary compared to full time teachers with a BA?

2. Set up the problem as a hypothesis testing problem. State the Null, the Alternative, and the event E
Null -  full time teachers with an MA start with a no significantly higher minimum base salary compared to full time teachers with a BA.
Alternative -  full time teachers with an MA start with a no significantly higher minimum base salary compared to full time teachers with a BA.
E - Test - A good test is measured by the value:
\[ D = \sum_{i} |d_i - \overline{d}|  \]
where $d_i$ are the  higher minimum base salary between full time teachers by degree. This measures the deviation from the population average. $p$-value: Using the test statistic above, the $p$-value is:\[ p  = \mathbb{P}(D \geq d) \] - We can use the permutation test to compute an approximation to $p$. Then use $p$ to test Null hypothesis

3. Compute an estimation of the p-value using the *permutation test*. (NO other tests accepted). 


```{r}
base.shuffle.min <- function(){
  fake.base <- sample(fulltime$base)
  fake.BA.min <- min(fake.base[fulltime$degree=="BA"])
  fake.MA.min <- min(fake.base[fulltime$degree=="MA"])
  fake.diff <- abs(fake.BA.min-fake.MA.min)
  return(fake.diff)
}
#replicate it many times
S.min <- replicate(10**4,base.shuffle.min())


```

```{r}
abs.diff.min <- abs(min(fulltime$base[fulltime$degree=="BA"])-min(fulltime$base[fulltime$degree=="MA"]))
p.hat.min <- sum(S.min>=abs.diff.min)/length(S.min)
p.hat.min
```

4. Draw conclusions. 
The calculated p value is small, although we should not reject $H_0$ completly. but there is a somewhat difference in full time teachers with an MA start with a no significantly higher minimum base salary compared to full time teachers with a BA. 


### Is there a significant difference in the *standard deviation* of base salary between *full time* teachers with BA and teachers with MA?

2. Set up the problem as a hypothesis testing problem. State the Null, the Alternative, and the event E

Null -  full time teachers with an MA start with a no significantly in the standard deviation base salary compared to full time teachers with a BA.
Alternative -  full time teachers with an MA start with a no significantly standard deviation base salary compared to full time teachers with a BA.
Test - A good test is measured by the value:
\[ D = \sum_{i} |d_i - \overline{d}|  \]
where $d_i$ are the standard deviation base salary between full time teachers by degree. This measures the deviation from the population average. $p$-value: Using the test statistic above, the $p$-value is:\[ p  = \mathbb{P}(D \geq d) \] - We can use the permutation test to compute an approximation to $p$. Then use $p$ to test Null hypothesis.

3. Compute an estimation of the p-value using the *permutation test*. (NO other tests accepted). 


```{r}
base.shuffle.sd <- function(){
  fake.base <- sample(fulltime$base)
  fake.BA.sd <- sd(fake.base[fulltime$degree=="BA"])
  fake.MA.sd <- sd(fake.base[fulltime$degree=="MA"])
  fake.diff <- abs(fake.BA.sd-fake.MA.sd)
  return(fake.diff)
}
#replicate it many times
S.sd <- replicate(10**4,base.shuffle.sd())

```

```{r}
abs.diff.sd <- abs(sd(fulltime$base[fulltime$degree=="BA"])-sd(fulltime$base[fulltime$degree=="MA"]))
p.hat.sd <- sum(S.sd>=abs.diff.sd)/length(S.sd)
p.hat.sd
```
4. Draw conclusions. 
The calculated p value is pretty high, although we should not agree on $H_0$ completly. but there is a little to no difference in full time teachers with an MA start with a no significantly higher minimum base salary compared to full time teachers with a BA.
## Sampling 

Do exercise 1.42 in the OpenIntro textbook.

Screens, teens, and psychological well-being. In a study of three nationally representative large-
scale data sets from Ireland, the United States, and the United Kingdom (n = 17,247), teenagers between
the ages of 12 to 15 were asked to keep a diary of their screen time and answer questions about how they
felt or acted. The answers to these questions were then used to compute a psychological well-being score.
Additional data were collected and included in the analysis, such as each child's sex and age, and on the
mother's education, ethnicity, psychological distress, and employment. The study concluded that there is
little clear-cut evidence that screen time decreases adolescent well-being.35


(a) What type of study is this?
Observation

(b) Identify the explanatory variables.

teenagers between the ages of 12 to 15 who watch TV daily.

(c) Identify the response variable.

adolescent well-being

(d) Comment on whether the results of the study can be generalized to the population, and why.
The study can be generalized to the population since it tested on 3 different relativelylarge population countries.

(e) Comment on whether the results of the study can be used to establish causal relationships.

The study can be used to establish casual relationship since screen time can be affect by the mother's education, ethnicity, psychological distress, and employment.

1.43 Stanford Open Policing. The Stanford Open Policing project gathers, analyzes, and releases records