title: "Homework 1 - Descriptive Statistics"
date: "Due date: Tuesday September 10, 2019"
output: pdf_document
```{r setup, include=FALSE}knitr::opts_chunk$set(echo = TRUE)```

1.	What is the rate of pairs who want to stay as partners?  Does this rate differ by gender pairing? (male/male, male/female vs female/male , ).

This show male and male pairing table.Where is 0 is didn't want to pair and 1 is want to pair again.

```{r}
setwd("/Users/williamkwon/Documents/R project folder")
data <- read.csv("speedPartner.csv", header = T)
data$PersonID <- rep(c(1,2), dim(data)[1]/2)
date.df <- reshape(data, idvar = "PairID", timevar = "PersonID", direction = "wide")
select <- date.df$Gender.1 == "Male" & date.df$Gender.2 == "Male"
date.df.mm <- date.df[select,]
date.df.mm$outcome <- date.df.mm$Decision.1*date.df.mm$Decision.2
table(date.df.mm$outcome)
```

This show male and femal pairing table.

```{r}
select <- date.df$Gender.1 == "Female" & date.df$Gender.2 == "Female"
date.df.ff <- date.df[select,]
date.df.ff$outcome <- date.df.ff$Decision.1*date.df.ff$Decision.2
table(date.df.ff$outcome)
```

This shows male and female paring table

```{r}
select <- date.df$Gender.1 == "Female" & date.df$Gender.2 == "Male" | date.df$Gender.1 == "Male" & date.df$Gender.2 == "Female"
sum(select)
date.df.mf <- date.df[select,]
date.df.mf$outcome <- date.df.mf$Decision.1*date.df.mf$Decision.2
table(date.df.mf$outcome)
```

If female and male or male and female then 68.75% want to stay as pairs.If they are male and male 83.33% want to stick together.If they are female and female they want to stay 100%.

2.	How well can people predict if the partner wants to stay with them? Is there a difference by gender?If female and female they want to pair again more than any other genders pair and male and male and female have the least desire to pair again.




```{r}
data$PersonID <- rep(c(1,2), dim(data)[1]/2)
data$Gender <- as.character(data$Gender)
date.df <- reshape(data, idvar = "PairID", timevar = "PersonID", direction = "wide")
df1 <- date.df[,c("Gender.1", "PartnerYes.1","Decision.2")]
df2 <- date.df[,c("Gender.2", "PartnerYes.2","Decision.1")]
names(df1) <- c("Gender", "PartnerYes" , "PartnerDecision")
names(df2) <- c("Gender", "PartnerYes" , "PartnerDecision")
pairagain <- rbind(df1,df2)
pairagain$Gender <-  as.factor(pairagain$Gender)
pairagain$PartnerYes.recode <- cut(date.df$PartnerYes.1, 3, labels=c("No"," Maybe", "Yes"))
mosaic(Gender ~PartnerYes.recode, data=pairagain)
```

So based on this graph they males are more likey to say they will predict pretty well but females are tend to unsure and put maybe.

1. Run the following two commands to produce two mosaic plots of Decision and Gender.

```{r}
mosaic(Decision ~ Gender, data = data)
```

```{r}
mosaic(Gender ~ Decision, data = data)
```


Thoes two plots are basically showing same conclusion but just flipped the axis.

Are people more likely to say ‘yes’ to people they like, or to those who they think likes them? Choose one or two most appropriate plots and/or table, and form your conclusion.

```{r}
mosaic(Decision ~ PartnerYes , data = data)
```

```{r}
mosaic(Decision ~ Like , data = data)
```

Based on Dicision vs PartnerYes and Decision vs Like People are more likely to say ‘yes’ to people they like than those who they think likes them.

1.	In one plot, make two side-by-side boxplots of spam vs exclaim_mess. 

```{r}
library(ggplot2)
library(vcd)
library(openintro)
data(email)
?email
email$spam <- as.factor(email$spam)
ggplot(email, aes(x=spam,y = exclaim_mess)) + geom_boxplot()
```

2. In one plot, make two overlaid histograms of spam vs exclaim_mess. 
```{r}
ggplot(email, aes(x=exclaim_mess, fill = spam)) + geom_histogram(binwidth=20)
```

3. What do these plots reveal? What are some problems with these plots?

Both plots reveal that less the The number of exclamation points in the email message would increase that email will be spam. The plot might have a problem since the numbers is too big so it’s hard to indicated than actual ratio of spam for exclaim_mess.


Variable recoding
Sometimes it is useful to recode variables. spam vs exclaim_mess is an example.

1.Recode exclaim_mess into four values:  0, 1, 2, >= 3.  You can write your own code, or use R’s cut
function, for example. Call this new variable exclaim_mess.recode.

```{r}
email$exclaim_mess.recode <- cut(email$exclaim_mess, breaks = c(-1,0,1,2,Inf),labels = c("less than 0","None","bigger than 1","At least 2"))
ggplot(email, aes(x=exclaim_mess.recode,fill=exclaim_mess)) + geom_bar()
```



2. What is the type of the variable exclaim_mess.recode?

```{r}
class(email$exclaim_mess.recode)
```
It's factor.

3. Produce a table and a mosaic plot of spam vs exclaim_mess.recode. What do they reveal?

```{r}
mosaic(spam ~ exclaim_mess.recode, data = email)
```

4.	Summarize the relationship between spam and exclaim_mess.recode in a couple of sentences.
More the exclaim_mess recode there are less chance to emails are spam.

